Old or Heavy?\\Decaying Gracefully with Age/Weight Shapes.

### 2019-02-27 papers/CADE19/paper.tex
fix results section

We set out to experimentally confirm two assumptions from folklore:
\begin{enumerate}
	\item The choice of age/weight ratio often has a significant effect on the performance of proof search
	\item There is no generally best age/weight ratio, even for a syntactically well-defined subset of problems
\end{enumerate}
These were supported by the work of \sandm{} but here we focus on these points and explore them more deeply.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.8\textwidth]{example-optimal-awr}
	\caption{
The number of active clauses reported by Vampire after successful 1-second runs on a TPTP problem.
For sufficiently negative \(x\), proof search is biased towards age to such an extent that a proof is found via breadth-first search (and hence the number of active clauses remains constant), whereas after \(x\) becomes sufficiently positive, proof search does not terminate for this problem in the given time.
In between these limits, the function is ``noisy'' and discontinuous.
These properties are typical for the observed set of problems.
}
	\label{fig:example-optimal-awr}
\end{figure}

@@

\begin{figure}[t]
	\centering
	\begin{tabular}{l l l l l}
AWR & Maximum & Mean & Standard Deviation\\
\hline
8:1 & 150.67 & 1.37 & 1.98\\
5:1 & 122.22 & 1.33 & 1.64\\
4:1 & 101.44 & 1.32 & 1.42\\
3:1 & 105.00 & 1.29 & 1.41\\
2:1 & 112.67 & 1.27 & 1.45\\
3:2 & 119.89 & 1.27 & 1.51\\
5:4 & 125.00 & 1.26 & 1.55\\
1 & 133.56 & 1.26 & 1.63\\
2:3 & 147.67 & 1.28 & 1.79\\
2 & 162.67 & 1.31 & 1.97\\
3 & 190.56 & 1.37 & 2.30\\
4 & 17.41 & 1.38 & 0.67\\
5 & 133.67 & 1.44 & 1.70\\
6 & 106.44 & 1.47 & 1.46\\
7 & 104.89 & 1.49 & 1.44\\
8 & 101.33 & 1.53 & 1.45\\
10 & 101.78 & 1.60 & 1.53\\
12 & 101.67 & 1.65 & 1.62\\
14 & 103.00 & 1.70 & 1.75\\
16 & 101.33 & 1.74 & 1.76\\
20 & 100.89 & 1.82 & 1.94\\
24 & 101.33 & 1.89 & 2.08\\
28 & 99.22 & 1.96 & 2.24\\
32 & 100.00 & 1.99 & 2.36\\
40 & 98.78 & 2.09 & 2.64\\
50 & 98.78 & 2.17 & 2.88\\
64 & 101.22 & 2.28 & 3.19\\
128 & 107.44 & 2.57 & 4.16\\
1024 & 222.01 & 2.83 & 7.55
	\end{tabular}
	\caption{The mean relative performance drop (in terms of activations compared to an optimal value) choosing a single AWR over an entire problem set causes. No setting is significantly ``better'' or ``worse'' than the others, but some have more variance in performance.}
	\label{fig:no-best-awr}
\end{figure}

@@

In experiments, several AWR ``shapes'' were commonly seen, including:
\begin{itemize}
	\item Near-constant, as seen currently.
	\item A decreasing trend from a higher AWR to a lower AWR.
	\item An increasing trend from a lower AWR to a higher AWR.
\end{itemize}
\todo{Need to be more specific about `commonly seen'}

@@

Together, these new settings allow a number of new option combinations, which can be used in conjunction with Vampire's portfolio \emph{CASC mode} pending integration into the strategy schedules.

@@

\subsection{Benchmark Performance}
The following shows the performance in terms of solved problems of all the configurations tested.
``Proved'' refers to the total number of problems a configuration solved, whereas ``Uniques'' is the number of problems a configuration solved which were not solved by the baseline.
These data show that configurations which are more similar to the baseline (i.e. slow decay or fast convergence) achieve more similar performance, as expected.
In total, 134 problems were solved by the new configurations that were not solved by the baseline.

\begin{center}
\begin{tabular}{l l l l}
	Configuration & Frequency & Proved & Uniques\\
	\hline
	baseline & -- & 13057 & 0\\
	converge & 1 & 13039 & 24\\
	converge & 5 & 13029 & 27\\
	converge & 10 & 13028 & 35\\
	converge & 50 & 13015 & 45\\
	converge & 100 & 12976 & 51\\
	converge & 500 & 12895 & 63\\
	converge & 1000 & 12837 & 52\\
	converge & 5000 & 12775 & 53\\
	converge & 10000 & 12751 & 53\\
	decay & 1 & 12698 & 48\\
	decay & 5 & 12702 & 51\\
	decay & 10 & 12698 & 48\\
	decay & 50 & 12712 & 49\\
	decay & 100 & 12726 & 46\\
	decay & 500 & 12795 & 29\\
	decay & 1000 & 12860 & 29\\
	decay & 5000 & 12982 & 16\\
	decay & 10000 & 13002 & 7\\
\end{tabular}
\end{center}

>>>>

We set out to experimentally confirm two assumptions from folklore:
\begin{enumerate}
	\item The choice of age/weight ratio often has a significant effect on the performance of proof search.
	\item There is in general no single best age/weight ratio for a given set of problems.
\end{enumerate}
These were supported by the work of \sandm{} but here we focus on these points and explore them more deeply.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=0.8\textwidth]{example-optimal-awr}
	\caption{
The number of active clauses reported by Vampire after successful 1-second runs on a TPTP problem.
In between the peaks on either side, the function of \(L\) is discontinuous with large peaks and troughs, but does follow an approximate trend and settles toward the global optimum.
These properties are typical for the observed set of problems.
}
	\label{fig:example-optimal-awr}
\end{figure}

@@

\begin{figure}[t]
	\centering
	\begin{tabular}{l l l l l}
AWR & Maximum & Mean & Standard Deviation\\
\hline
8:1 & 150.67 & 1.37 & 1.98\\
5:1 & 122.22 & 1.33 & 1.64\\
4:1 & 101.44 & 1.32 & 1.42\\
3:1 & 105.00 & 1.29 & 1.41\\
2:1 & 112.67 & 1.27 & 1.45\\
3:2 & 119.89 & 1.27 & 1.51\\
5:4 & 125.00 & 1.26 & 1.55\\
1 & 133.56 & 1.26 & 1.63\\
2:3 & 147.67 & 1.28 & 1.79\\
2 & 162.67 & 1.31 & 1.97\\
3 & 190.56 & 1.37 & 2.30\\
4 & 17.41 & 1.38 & 0.67\\
5 & 133.67 & 1.44 & 1.70\\
6 & 106.44 & 1.47 & 1.46\\
7 & 104.89 & 1.49 & 1.44\\
8 & 101.33 & 1.53 & 1.45\\
10 & 101.78 & 1.60 & 1.53\\
12 & 101.67 & 1.65 & 1.62\\
14 & 103.00 & 1.70 & 1.75\\
16 & 101.33 & 1.74 & 1.76\\
20 & 100.89 & 1.82 & 1.94\\
24 & 101.33 & 1.89 & 2.08\\
28 & 99.22 & 1.96 & 2.24\\
32 & 100.00 & 1.99 & 2.36\\
40 & 98.78 & 2.09 & 2.64\\
50 & 98.78 & 2.17 & 2.88\\
64 & 101.22 & 2.28 & 3.19\\
128 & 107.44 & 2.57 & 4.16\\
1024 & 222.01 & 2.83 & 7.55
	\end{tabular}
	\caption{The mean relative performance drop (in terms of activations compared to an optimal value) choosing a single AWR over an entire problem set causes. No setting is significantly ``better'' or ``worse'' than the others, but some have more variance in performance. A ratio of 1:4 produces an unusually well-behaved maximum performance drop. \sandm~found that 1:5 had a similar property, but this might be explained by differences in test environment.}
	\label{fig:no-best-awr}
\end{figure}

@@

Unsurprisingly, in ad-hoc experiments on other problems, the best shape is rarely constant.
This suggests that implementing other shapes, such as an increasing or decreasing trend, might lead to quicker proofs in the Vampire theorem prover.

@@

Together, these new settings allow a number of new option combinations, which can be used in conjunction with Vampire's portfolio \emph{CASC mode} pending integration into the strategy schedules.
This version of the prover is currently in a separate branch in the main Vampire repository\footnote{\texttt{https://github.com/vprover/vampire/tree/awr-shapes}}.

@@

\subsection{Benchmark Performance}
Figure \ref{fig:results} shows the performance in terms of solved problems of all the configurations tested.
These data show that configurations which are more similar to the baseline (i.e. slow decay or fast convergence) achieve more similar performance, as expected.
In total, 134 problems were solved by the new configurations that were not solved by the baseline.

\begin{figure}
\centering
\begin{tabular}{l l l l l l}
	Configuration & Frequency & Proved & Fresh & Uniques & \(u\)-score\\
	\hline
	baseline & -- & 13057 & 0 & 1 & 714.2\\
	converge & 1 & 13039 & 24 & 3 & 714.3\\
	converge & 5 & 13029 & 27 & 1 & 709.5\\
	converge & 10 & 13028 & 35 & 5 & 714.3\\
	converge & 50 & 13015 & 45 & 5 & 712.8\\
	converge & 100 & 12976 & 51 & 1 & 705.9\\
	converge & 500 & 12895 & 63 & 4 & 698.3\\
	converge & 1000 & 12837 & 52 & 0 & 688.6\\
	converge & 5000 & 12775 & 53 & 1 & 682.4\\
	converge & 10000 & 12751 & 53 & 0 & 678.7\\
	decay & 1 & 12698 & 48 & 1 & 673.6\\
	decay & 5 & 12702 & 51 & 1 & 674.9\\
	decay & 10 & 12698 & 48 & 1 & 674.2\\
	decay & 50 & 12712 & 49 & 2 & 679.1\\
	decay & 100 & 12726 & 46 & 1 & 678.8\\
	decay & 500 & 12795 & 29 & 1 & 685.5\\
	decay & 1000 & 12860 & 29 & 2 & 692.6\\
	decay & 5000 & 12982 & 16 & 2 & 707.1\\
	decay & 10000 & 13002 & 7 & 0 & 706.3\\
\end{tabular}
\caption{
Results for the tested configurations.
\emph{Proved} refers to the total number of problems a configuration solved.
\emph{Fresh} is the number of problems a configuration solved which were not solved by the baseline.
\emph{Uniques} is the number of problems a configuration solved which were not solved by any other configuration.
\emph{\(u\)-score} is a refined unique score which correlates to a configuration's utility in solving new problems, as used in Hoder \emph{et al.}~\cite{u-score}.
}
\label{fig:results}
\end{figure}
